---
title: "Personal Movement Monitoring"
author: "Kurt C Anderson"
date: "March 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This report makes use of data from a personal activity monitoring device. The devices collect data at 5 minute intervals throughout the day. The data consists of two months collected from anonymous individuals, during the months of October and November, 2012. 
Each observation includes the number of steps taken in 5 minute intervals each day.

## Loading and preprocessing the data
### Download and unzip dataset
```{r echo=TRUE}
## setwd("~/Documents/R Projects/Coursera/reproducible research/Project1")
if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl,destfile="./data/Dataset.zip")

# Unzip dataSet to /data directory
unzip(zipfile="./data/Dataset.zip",exdir="./data")

```

### Read data
#### This first data frame contains all rows including those missing values (NA)
#### The count of rows is below
```{r echo=TRUE}
a_data <-  read.csv("./data/activity.csv", header=TRUE)
dim(a_data)
```

## What is mean total number of steps taken per day?
### Summary statistics including Mean and Median
```{r echo=TRUE}
summary(a_data$steps)
```

### Histogram of the total steps taken each day
```{r echo=TRUE}
hist(a_data$steps,
     main="Histogram for Steps Taken")
```

## Imputing missing values
### Copy data to a new DF removing NAs
#### This is a copy of previous data frame with the NAs row removed
#### The new count of rows is below
```{r echo=TRUE}
activity_data <- na.omit(a_data)
dim(activity_data)
```

## Are there differences in activity patterns between weekdays and weekends?